Approximation as Differentiation

### 2025-07-02 main/approx-as-tangents.tex
Cite Dex, Sigal, etc.

We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a denotational setting. From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps\bob{and we did it in Agda, so we can run it}.

@@

Derivatives can be computed numerically by computing $f$ on small perturbations of its input, or symbolically by examining a closed-form representation of $f$. However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output. \bob{cite something on autodiff}. This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms~\cite{jax2018github}. \bob{cite: TensorFlow, Dex, Jesse Sigal, etc.}

>>>>

We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a denotational setting. From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps. \bob{and we did it in Agda, so we can run it}

@@

Derivatives can be computed numerically by computing $f$ on small perturbations of its input, or symbolically by examining a closed-form representation of $f$. However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output~\cite{linnainmaa76}. This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms~\cite{jax2018github,abadi16,elliott17,sigal24}.
